Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2023-02-17 analysis/paper/paper.qmd
Finish editing the text

Overall, our results show a wide range of TCSA values throughout the Late Paleolithic period and between eco-regional zones. We confirmed the widest TCSA range at a single site, Suyanggae (@fig-tcsa-per-sites). Our foundings are therefore consistent with our second scenario, in which stemmed points are described as multifunctional tools, suggesting that people created stemmed points as a response to unexpected or varying circumstances in their habitats, similar to how @eren2022plains explain Clovis points.

@@

Considering the importance of stemmed points for the technological transition during the Late Paleolithic in Korea, our research examines the likely uses for stemmed points by asking three research questions: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We applied the tip cross-sectional area (TCSA) metric because it has been used as a ballistically relevant standard to discriminate different likely use classes of projectile points and it requires only a few measurements on a stone artifact (i.e. width and thickness) to compute the metric. We calculated TCSA for a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites. Then we examined the TCSA values with other variables including raw materials, weight, radiocarbon dates, and site locations. Drawing on evolutionary theory, we premised that the way of using stemmed points is likely to reflect the socio-environmental circumstances that people encountered and managed. We examined the possible impact of LGM on stemmed points and their distribution across eco-regional zones.

According to the different weapon-delivery systems that can be inferred from TCSA values, the majority of stemmed points from the Korean Palaeolithic were probably used as javelin tips and stabbing spear tips. In general, though, we noted a wide range and also differing distributions of TCSA values in each assemblage. Therefore, we conclude that stemmed points had diverse roles during the Late Paleolithic. We conclude that people frequently encountered unfamiliar situations and used stemmed points as a multifunctional tool to carry out multiple tasks. We are aware that discriminating the likely use of small numbers of projectile points could be arbitrary [@erlandson2014darts]. Since TCSA covers the critical elements of projectiles, flight and penetration dynamics (i.e. increased or decrease by shape of tip and cross section), we still consider it a useful metric for discriminating between weapon tip tools [@hughes1998getting; @sitton2020tip; @lombard2021variation]. As part of our future research, we will conduct use-wear analyses of stemmed points and experimental research in an effort to elucidate the detailed function of stemmed points and technological links between hunter-gatherer groups.

>>>>

Overall, our results show a wide range of TCSA values throughout the Late Paleolithic period and between eco-regional zones. We confirmed the widest TCSA range at a single site, Suyanggae (@fig-tcsa-per-sites). Our findings are therefore consistent with our second scenario, in which stemmed points are described as multifunctional tools, suggesting that people created stemmed points as a response to unexpected or varying circumstances in their habitats, similar to how @eren2022plains explain Clovis points.

@@

Considering the importance of stemmed points for the technological transition during the Late Paleolithic in Korea, our research examines the likely uses for stemmed points by asking three research questions: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We applied the tip cross-sectional area (TCSA) metric because it has been used as a ballistically relevant standard to discriminate different likely use classes of projectile points and it requires only a few measurements on a stone artifact (i.e. width and thickness) to compute the metric. We calculated TCSA for a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites. Then we examined the TCSA values with other variables including raw materials, weight, radiocarbon dates, and site locations. Drawing on evolutionary theory, we premised that the way of using stemmed points are likely to reflect the socio-environmental circumstances that people encountered and managed. We examined the possible impact of LGM on stemmed points and their distribution across eco-regional zones.

According to the different weapon-delivery systems that can be inferred from TCSA values, the majority of stemmed points from the Korean Palaeolithic were probably used as javelin tips and stabbing spear tips. In general, though, we noted a wide range and also differing distributions of TCSA values in each assemblage. Therefore, we conclude that stemmed points served diverse roles during the Late Paleolithic. People might frequently encounter unfamiliar situations and use stemmed points as a multifunctional tool to carry out multiple tasks. We are aware that discriminating the likely use of small numbers of projectile points could be arbitrary [@erlandson2014darts]. Since TCSA covers the critical elements of projectiles, flight and penetration dynamics (i.e. increase or decrease by shape of tip and cross section), we still consider it a useful metric for discriminating between weapon tip tools [@hughes1998getting; @sitton2020tip; @lombard2021variation]. As part of our future research, we will conduct use-wear analyses of stemmed points and experimental research in an effort to elucidate the detailed function of stemmed points and technological links between hunter-gatherer groups.
